Lab 2 Draft Report: COVID-19 Case Rate vs Population and Policy

### 2020-12-03 Jackson_Liu_Temlock_Zhang_Lab2_Draft_Report.Rmd
Update Jackson_Liu_Temlock_Zhang_Lab2_Draft_Report.Rmd
update cook distance for model3
model_2 update, 2-a, 2-b deleted

@@

### Model 2 Regression
The regression models regarding the demographic variables discussed previously are listed as below.However, no coefficients regarding these variables in model 2 are statistically significant.

```{r}
df<-df %>%
  mutate(
    noneld_pre_existing_per_100k = 100000*non_elderly_pre_existing/population_2018,
    white_per_100k = 100000*white_population_pct,
    senior_per_100k = 100000*senior_pct,
    poverty_per_100k = 100000*poverty_pct, 
    unemployed_per_100k = 100000*unemployed_pct,
    log_population_density = log(population_density),
    avg_distance = sqrt(2589988.11/population_density),
    area_per_capita = 2589911.11/population_density
  )
```

#### Model 2-a

$$
\begin{aligned}
 Case.Rate.Per.100K =\beta_0 &+\beta_1Senior.Rate.Per.100K \\&+\beta_2Nonelderly.Pre.Existing.Rate.Per.100K\\&+\beta_3Poverty.Rate.Per.100K+\beta_4ln(Population.Density)\\&+\beta_5White.Rate.Per.100K
\end{aligned}
$$

```{r}
model2a <- lm(case_rate_100k ~ senior_per_100k + log(population_density) + noneld_pre_existing_per_100k + poverty_per_100k + white_per_100k, data = df)
coeftest(model2a)
```

#### Model 2-b

$$
\begin{aligned}
 Case.Rate.Per.100K =\beta_0 &+\beta_1Senior.Rate.Per.100K \\&+\beta_2Nonelderly.Pre.Existing.Rate.Per.100K\\&+\beta_3Poverty.Rate.Per.100K+\beta_4ln(Population.Density)\\&+\beta_5White.Rate.Per.100K+\beta_6Unemployed.Per.100k
\end{aligned}
$$
```{r}
model2b <- lm(case_rate_100k ~ senior_per_100k + log(population_density) + noneld_pre_existing_per_100k + poverty_per_100k + white_per_100k + unemployed_per_100k, data = df)
coeftest(model2b)
```

#### need to update
From the results of the regression, none of the coefficients of investigated variables in model 2 are significant (p-value < 0.05). Among these regression coefficient tests,  the p-value regarding the poverty percentage variable is the lowest (0.05755).  In addition, its positive estimated coefficient is consistent with the statement that low-income communities are at high pandemic risk claimed by mainstream media these days.

@@

@@

#### 2-a
#### i. IID Sampling 
( Discussed in the common section)

@@

```{r, echo=FALSE}
df<- df%>% 
  mutate(
    model2a_preds = predict(model2a), 
    model2a_resids = resid(model2a)
  ) 
df %>% 
  ggplot(aes(model2a_preds, model2a_resids)) + 
  geom_point() + 
  stat_smooth()
```

@@

```{r}
model2a$coefficients
```

@@

```{r}
vif(model2a)
```

####iv.Homoskedastic Errors
32 is New Mexico, 33 is New York, 35 is North Dakota

```{r, echo=FALSE}
plot(model2a, which=3)
```

@@

```{r, echo=FALSE}
plot_one <- df %>% 
  ggplot(aes(x = model2a_resids)) + 
  stat_bin(bins = 25)
  geom_histogram()

plot_two <- df %>% 
  ggplot(aes(sample = model2a_resids)) + 
  stat_qq() + stat_qq_line()

@@

```{r, echo=FALSE}
df %>% 
  select(case_rate_100k, senior_per_100k, log_population_density, noneld_pre_existing_per_100k, poverty_per_100k, white_per_100k) %>% 
  GGally::ggpairs()
```

#### Cook Distance 
9-DC, 12-Hawaii, 32-New Mexico.

```{r, echo =FALSE}
plot(model2a, which = 5)
```

#### 2-b
####i.IID Sampling

#####ii.Linear Conditional Expectation
Assessing non-linear in higher-dimensional space: look at the predicted vs. residuals of the model.

```{r, echo=FALSE}
df<- df%>% 
  mutate(
    model2b_preds = predict(model2b), 
    model2b_resids = resid(model2b)
  ) 
df %>% 
  ggplot(aes(model2b_preds, model2b_resids)) + 
  geom_point() + 
  stat_smooth()
```

###iii.No Perfect Collinearity
First, check if any variables were dropped by R. No coefficient is missing, which mean there is no perfect collinearity.

```{r}
model2b$coefficients
```

```{r}
vif(model2b)
```

###iv.Homoskedastic Errors
32 is New Mexico, 33 is New York, 35 is North Dakota

```{r, echo=FALSE}
plot(model2b, which=3)
```

###v.Normally Distributed Errors
Check the normality of error distribution based on the histogram of residuals and the qqplot.

```{r, echo=FALSE}
plot_one <- df %>% 
  ggplot(aes(x = model2b_resids)) + 
  stat_bin(bins = 25)
  geom_histogram()

plot_two <- df %>% 
  ggplot(aes(sample = model2b_resids)) + 
  stat_qq() + stat_qq_line()

#### GGPairs
```{r, echo=FALSE}
df %>% 
  select(case_rate_100k, senior_per_100k, log_population_density, noneld_pre_existing_per_100k, poverty_per_100k, white_per_100k) %>% 
  GGally::ggpairs()
```

### Cook Distance 
9-DC, 12-Hawaii, 32-New Mexico.

```{r, echo =FALSE}
plot(model2b, which = 5)
```

@@

```{r warning=FALSE}
se.model1 = coeftest(model1, vcov = vcovHC)[ , "Std. Error"]
se.model2 = coeftest(model2b, vcov = vcovHC)[ , "Std. Error"]
se.model3 = coeftest(model3a, vcov = vcovHC)[ , "Std. Error"]

stargazer(model1, model2b, model3a, type = "text", omit.stat = "f",
          se = list(se.model3),
          star.cutoffs = c(0.05, 0.01, 0.001), 
          title = "Table 1: The relationship between COVID case rate and population density")

>>>>

# percentage varialbe re-scale to number_per_100k (need to double check where to put this part in EDA or model regression)

df<-df %>%
  mutate(
    noneld_pre_existing_per_100k = 100000*non_elderly_pre_existing/population_2018,
    white_per_100k = 100000*white_population_pct,
    senior_per_100k = 100000*senior_pct,
    poverty_per_100k = 100000*poverty_pct, 
    unemployed_per_100k = 100000*unemployed_pct,
    log_population_density = log(population_density),
    avg_distance = sqrt(2589988.11/population_density),
    area_per_capita = 2589911.11/population_density
  )
```

@@

### Model 2 Regression
The regression models regarding the demographic variables discussed previously are listed as below.

$$
\begin{aligned}
 Case.Rate.Per.100K =\beta_0 &+\beta_1Senior.Rate.Per.100K +\beta_2Poverty.Rate.Per.100K +\beta_3Unemployed.Per.100K
\end{aligned}
$$

```{r}
model2 <- lm(case_rate_100k ~ senior_per_100k + poverty_per_100k + unemployed_per_100k, data = df)
coeftest(model2)
```

#### (need to update)
From the results of the regression, none of the coefficients of investigated variables in model 2 are significant (p-value < 0.05). Among these regression coefficient tests,  the p-value regarding the poverty percentage variable is the lowest (0.05755).  In addition, its positive estimated coefficient is consistent with the statement that low-income communities are at high pandemic risk claimed by mainstream media these days.

@@

#Cook distance
plot(model3a, which=5)

@@

#### i. IID Sampling 
( Discussed in the common section)

@@

```{r, echo=FALSE}
df<- df%>% 
  mutate(
    model2_preds = predict(model2), 
    model2_resids = resid(model2)
  ) 
df %>% 
  ggplot(aes(model2_preds, model2_resids)) + 
  geom_point() + 
  stat_smooth()
```

@@

```{r}
model2$coefficients
```

@@

```{r}
vif(model2)
```

####iv.Homoskedastic Errors
10 is Florida, 33 is New York, 35 is North Dakota

```{r, echo=FALSE}
plot(model2, which=3)
```

@@

```{r, echo=FALSE, warninng=False}
plot_one <- df %>% 
  ggplot(aes(x = model2_resids)) + 
  stat_bin(bins = 25)
  geom_histogram()

plot_two <- df %>% 
  ggplot(aes(sample = model2_resids)) + 
  stat_qq() + stat_qq_line()

@@

```{r, echo=FALSE}
df %>% 
  select(case_rate_100k, senior_per_100k, poverty_per_100k, unemployed_per_100k) %>% 
  GGally::ggpairs()
```

#### Cook Distance 
10 Florida, 20 Maine, 35 North Dakota

```{r, echo =FALSE}
plot(model2, which = 5)
```

@@

```{r warning=FALSE}
se.model1 = coeftest(model1, vcov = vcovHC)[ , "Std. Error"]
se.model2 = coeftest(model2, vcov = vcovHC)[ , "Std. Error"]
se.model3 = coeftest(model3a, vcov = vcovHC)[ , "Std. Error"]

stargazer(model1, model2, model3a, type = "text", omit.stat = "f",
          se = list(se.model3),
          star.cutoffs = c(0.05, 0.01, 0.001), 
          title = "Table 1: The relationship between COVID case rate and population density")
